Fail Faster

### 2025-03-19 paper/paper.tex
minor

\jwc{Initial section here about eval}

\begin{itemize}
  \item \textbf{RQ1}: Do generators written using \name run faster than those written with regular generator combinators?
  \item \textbf{RQ2}: Do performance speedups translate to better bug-finding ability?
  % \item \textbf{RQ3}: Does our technique generalize to other strict functional languages?
\end{itemize}

@@

\subsection{Benchmarking Generator Speed}
\jwc{Intro stuff}

\begin{figure}[h]
  \centering
  \vspace{0.5em} % Kept the vertical spacing
  \begin{tabular}{lrl}
  \toprule
  Benchmark & \multicolumn{2}{c}{avg. time (ns/op) $\pm$ std. dev.} \\
  \midrule
  BST & $3,560.813$ & $\pm\>1,511.201$ \\
  BST, Staged & $751.287$ & $\pm\>39.890$ \\
  BST (Type-Derived) & $44,619.863$ & $\pm\>5,761.568$ \\
  BST (Type-Derived), Staged & $6,863.209$ & $\pm\>416.188$ \\
  RBT (Type-Derived) & $60,559.846$ & $\pm\>4,901.980$ \\
  RBT (Type-Derived), Staged & $8,511.528$ & $\pm\>324.615$ \\
  STLC & $44,497.623$ & $\pm\>1,970.484$ \\
  STLC, Staged & $7,151.259$ & $\pm\>355.639$ \\
  \bottomrule
  \end{tabular}
 \caption{\todo{Understand Joe's implementation of this...}}
 \label{fig:benchmark-table}
 \end{figure}

\begin{figure}[h]
  % \includegraphics[scale=0.33]{allocs.png}
  \includegraphics[scale=0.45]{time.png}
  \label{fig:benchmark-time}
  \caption{Average time taken to generate a value with the staged generator, in nanoseconds.}
\end{figure}
\jwc{NOTE: we should test generator speed across both languages, but speed -> bugfinding ability in only OCaml.}
\jwc{
  Baseline generators to test speed in both languages:
  \begin{itemize}
    \item Single int
    \item Pair of ints, constrained
    \item List of ints without bind (use map): lots of sampling, minimal binds.
    \item Unableled Trees of a fixed size (no weighted union) minimal sampling, lots of binds.
  \end{itemize}
}
\subsection{Benchmarking Bug-Finding Ability}
To answer \textbf{RQ2}, we evaluate the bug-finding ability of our BST and STLC case studies using Etna, a platform for assessing and comparing PBT techniques. 
Etna allows users to measure the effectiveness of different generator implementations by injecting bugs into the system under test and recording the time taken for a relevant 
property to fail in response. For example, in a tree with an ordering property, a bug injection might consist of inserting a value into the tree that violates that ordering. Each case study includes a diverse set of \textit{tasks}, where a task consists of a specific bug-property pair designed to test different aspects of the system (e.g., BST
includes tasks that test insertion, deletion, and union operations on binary search trees).

@@

It is worth noting that the time-to-failure of a given strategy on a given seed is not necessarily representative of 
the strategy's average time-to-failure over a large number of trials. We normalize by computing the
timing \textit{differential} between optimized and unoptimized generators (the ``speedup''). While this works in most cases,
it is not perfect.
For example, it is theoretically possible
to choose a seed such that the \textit{first} value produced by a generator discovers the bug, which would elide any differential speedup (and likewise
if both generators fail to find the bug).
To account for these edge cases, we repeat the above process for 10 seeds and compute the geometric mean of the speedups.\footnote{Although the variance \textit{between} seeds in Etna can be vast, timing results are replicable for a \textit{given} seed. Across 1,000 trials using the same seed, 
the observed variance in time-to-failure was less than a nanosecond.}

\begin{figure}[h]
  \includegraphics[scale=0.333]{speedup.png}
  \caption{Geometric average of each task's speedup over case study and strategies, showing that staging leads to better bug-finding ability across the board. As in \todo{earlier figure}, benchmarks with higher numbers of binds (BST (Single-Pass), STLC) see more benefit from staging, whereas benchmarks with more sampling (BST) see a greater speedup from the CSR \rand.}
  \label{fig:etna}
\end{figure}
\begin{figure}[h]
  \includegraphics[scale=0.355]{hist.png}
  \label{fig:swarm}
  \caption{Histogram of individual-task speedups across different staging strategies and benchmarks. Extreme outliers correspond to large speedups.}
\end{figure}

>>>>

\jwc{Initial section here about eval}\tr{todo}

\begin{itemize}
  \item \textbf{RQ1}: Do generators written using our technique run faster than those written with regular generator combinators?
  \item \textbf{RQ2}: Do performance speedups translate to better bug-finding ability?
  % \item \textbf{RQ3}: Does our technique generalize to other strict functional languages?
\end{itemize}

@@

\subsection{Benchmarking Generator Speed}

\begin{figure}[h]
  % \includegraphics[scale=0.33]{allocs.png}
  \includegraphics[scale=0.45]{time.png}
  \label{fig:time}
  \caption{Average generation time over 10,000 trials. Across all benchmarks, we see reliable improvement from at least one of Staged SR and Staged CSR.}
\end{figure}

\begin{figure}[h]
  \centering
  \vspace{0.5em} % Kept the vertical spacing
  \begin{tabular}{lrl}
  \toprule
  Benchmark & \multicolumn{2}{c}{avg. time (ns/op) $\pm$ std. dev.} \\
  \midrule
  BST & $3,560.813$ & $\pm\>1,511.201$ \\
  BST, Staged & $751.287$ & $\pm\>39.890$ \\
  BST (Type-Derived) & $44,619.863$ & $\pm\>5,761.568$ \\
  BST (Type-Derived), Staged & $6,863.209$ & $\pm\>416.188$ \\
  RBT (Type-Derived) & $60,559.846$ & $\pm\>4,901.980$ \\
  RBT (Type-Derived), Staged & $8,511.528$ & $\pm\>324.615$ \\
  STLC & $44,497.623$ & $\pm\>1,970.484$ \\
  STLC, Staged & $7,151.259$ & $\pm\>355.639$ \\
  \bottomrule
  \end{tabular}
 \caption{Benchmarks reimplemented in ScalaCheck show similar performance improvements to \name.}
 \label{fig:scala}
 \end{figure}

%\jwc{NOTE: we should test generator speed across both languages, but speed -> bugfinding ability in only OCaml.}
%\jwc{
%  Baseline generators to test speed in both languages:
%  \begin{itemize}
%    \item Single int
%    \item Pair of ints, constrained
%    \item List of ints without bind (use map): lots of sampling, minimal binds.
%    \item Unableled Trees of a fixed size (no weighted union) minimal sampling, lots of binds.
%  \end{itemize}
%}
\subsection{Benchmarking Bug-Finding Ability}
To answer \textbf{RQ2}, we evaluate the bug-finding ability of our BST and STLC case studies using Etna, a platform for assessing and comparing PBT techniques. 
Etna allows users to measure the effectiveness of different generator implementations by injecting bugs into the system under test and recording the time taken for a relevant 
property to fail in response. For example, in a tree with an ordering property, a bug injection might consist of inserting a value into the tree that violates that ordering. Each case study includes a diverse set of \textit{tasks}, where a task consists of a specific bug-property pair designed to test different aspects of the system (e.g., BST
includes tasks that test insertion, deletion, and union operations on binary search trees). BST has 43 tasks; STLC has 18.

@@

It is worth noting that the time-to-failure of a given strategy on a given seed is not necessarily representative of 
the strategy's average time-to-failure over a large number of trials. We normalize by computing the
timing \textit{differential} between optimized and unoptimized generators (the ``speedup''). While this works in most cases,
it is not perfect.
For example, it is theoretically possible
to choose a seed such that the \textit{first} value produced by a generator discovers the bug, which would elide any differential speedup (and likewise
if both generators fail to find the bug).
To account for these edge cases, we repeat the above process for 10 seeds and compute the geometric average of the speedups.\footnote{Although the variance \textit{between} seeds in Etna can be vast, timing results are replicable for a \textit{given} seed. Across 1,000 trials using the same seed, 
the observed variance in time-to-failure was less than a nanosecond.}

\begin{figure}[h]
  \includegraphics[scale=0.333]{speedup.png}
  \caption{Geometric average of each task's speedup over case studies and strategies, showing that staging leads to better bug-finding ability across the board. As in \todo{earlier figure}, benchmarks with higher numbers of binds (BST (Single-Pass), STLC) see more benefit from staging, whereas benchmarks with more sampling (BST) see a greater speedup from the CSR \rand.}
  \label{fig:etna}
\end{figure}

\begin{figure}[h]
  \includegraphics[scale=0.355]{hist.png}
  \label{fig:swarm}
  \caption{Histogram of individual-task speedups across different staging strategies and benchmarks. Extreme outliers correspond to large speedups.}
\end{figure}
